---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file. -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# minex

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/DIGlabUAB/minex/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/DIGlabUAB/minex/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`reprex` makes your bug **reproducible**. `minex` makes it **minimal**.

When you ask for help with R, you are told to post a minimal reproducible
example. The `reprex` package runs and formats your code; it does not shrink it.
`minex` shrinks it: paste in a failing script and it returns a one-minimal set of
statements that still throws the same error, finding it by delta debugging
rather than by deleting lines and re-running by hand. One-minimal means no
remaining statement can be removed individually while preserving the failure;
it does not claim the globally smallest possible script.

## Installation

```r
# install.packages("pak")
pak::pak("DIGlabUAB/minex")
```

## Usage

```{r}
library(minex)

script <- c(
  "a <- 10",
  "b <- 20",
  "log('not a number')"
)

minex(code = script)
```

The setup lines are dropped and only the offending statement remains. A
statement that the failure genuinely depends on is kept, because removing it
changes the error and the reduction is rejected.

`minex()` also reads a file directly:

```r
minex(file = "analysis.R")
```

And `reduce_rows()` does the same job for a data frame, returning the rows that
still reproduce a failure:

```{r}
df <- data.frame(id = 1:6, value = c(3, 8, 999, 2, 5, 7))
reduce_rows(df, function(d) any(d$value > 100))
```

## How it works

Each candidate reduction runs in a fresh R process (via `callr`) so that
statement dependencies and side effects are respected. A candidate "reproduces"
the failure when it errors with the same message as the original; matching on
the condition class instead, or supplying a custom oracle, is also supported.
The engine is `ddmin()`. Its default strategy adapts the counter-based block
sizing of [Zhang et al. (2025)](https://doi.org/10.1109/ICSE55347.2025.00117),
restarting the round counter after each successful reduction. A repeated
singleton-deletion sweep then ensures the result is one-minimal.
This sweep is related to, but is not the DDMIN* algorithm evaluated by
[Vince and Kiss (2024)](https://doi.org/10.1002/smr.2702). The classic algorithm
of
[Zeller and Hildebrandt (2002)](https://doi.org/10.1109/32.988498) remains
available with `algorithm = "ddmin"`.

If nothing can be removed at statement level, `minex()` automatically descends
into pipeline stages and positional call arguments using a specialized
HDD*-style pass
([Misherghi and Su, 2006](https://doi.org/10.1145/1134285.1134307)).

See `vignette("minex")` for the full walkthrough.
