---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file. -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# minex

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![R-CMD-check](https://github.com/DIGlabUAB/minex/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/DIGlabUAB/minex/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

`reprex` makes your bug **reproducible**. `minex` makes it **minimal**.

When you ask for help with R, you are told to post a minimal reproducible
example. The `reprex` package runs and formats your code; it does not shrink it.
`minex` shrinks it: paste in a failing script and it returns the smallest subset
of statements that still throws the same error, finding it by delta debugging
rather than by deleting lines and re-running by hand.

## Installation

```r
# install.packages("pak")
pak::pak("DIGlabUAB/minex")
```

## Usage

```{r}
library(minex)

script <- c(
  "a <- 10",
  "b <- 20",
  "log('not a number')"
)

minex(code = script)
```

The setup lines are dropped and only the offending statement remains. A
statement that the failure genuinely depends on is kept, because removing it
changes the error and the reduction is rejected.

`minex()` also reads a file directly:

```r
minex(file = "analysis.R")
```

And `reduce_rows()` does the same job for a data frame, returning the rows that
still reproduce a failure:

```{r}
df <- data.frame(id = 1:6, value = c(3, 8, 999, 2, 5, 7))
reduce_rows(df, function(d) any(d$value > 100))
```

## How it works

Each candidate reduction runs in a fresh R process (via `callr`) so that
statement dependencies and side effects are respected. A candidate "reproduces"
the failure when it errors with the same message as the original; matching on
the condition class instead, or supplying a custom oracle, is also supported.
The engine is `ddmin()`, an implementation of the delta debugging algorithm of
Zeller and Hildebrandt (2002), exposed for reuse on any collection.

See `vignette("minex")` for the full walkthrough.
