---
title: "Minimizing failing R code with minex"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Minimizing failing R code with minex}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(minex)
```

## The gap minex fills

When you ask for help with an R problem, the standard advice is to post a
*minimal reproducible example*. The
[reprex](https://cran.r-project.org/package=reprex) package handles the
*reproducible* half: it runs your snippet in a clean session and formats the
code together with its output. It does nothing about the *minimal* half. That
part, stripping a script down to the few lines that actually matter, is still
done by hand, and it is the tedious step.

`minex()` automates it. Give it failing code and it returns the smallest subset
of statements that still produces the same error.

## A first example

Here is a script where only the last line is to blame, surrounded by setup that
has nothing to do with the failure.

```{r}
script <- c(
  "a <- 10",
  "b <- 20",
  "log('not a number')"
)

res <- minex(code = script, backend = "inprocess")
res
```

The setup lines are gone. `as.character()` gives you the bare code, ready to
paste into a bug report.

```{r}
cat(as.character(res), "\n")
```

(The examples here use `backend = "inprocess"` so that they run quickly inside
the vignette. In normal use the default `backend = "callr"` evaluates each
candidate in a fresh R process, which is what you want when statements have side
effects.)

## Dependencies are respected

Reduction never throws away a statement the failure depends on. When a later
line needs an earlier one, removing the earlier line changes the error, and the
oracle rejects that reduction. Both lines below survive because they are jointly
required.

```{r}
script <- c(
  "x <- c(1, 2, NA)",
  "m <- mean(x)",
  "if (is.na(m)) stop('mean is NA')"
)
minex(code = script, backend = "inprocess")
```

## What counts as "the same failure"

By default a candidate must fail with the *same condition message* as the
original. This is usually the right choice: an over-reduced fragment tends to
fail differently (often "object not found"), and message matching rejects it.

If the message embeds changing details such as a value or an index, match on the
condition *class* instead:

```{r}
minex(code = script, match = "class", backend = "inprocess")
```

Use `match = "both"` to require the message and a shared class.

## Minimizing against an arbitrary condition

For full control, pass an `oracle`: a predicate over the statements that returns
`TRUE` when they still reproduce whatever you care about. With a custom oracle,
`minex()` does not record a target failure and `match` is ignored.

```{r}
minex(
  code = c("one <- 1", "two <- 2", "three <- 3"),
  oracle = function(stmts) any(grepl("two", stmts)),
  backend = "inprocess"
)
```

## Reducing the data, not just the code

A bug often hides in a large data frame even though a couple of rows are enough
to trigger it. `reduce_rows()` runs the same search over rows.

```{r}
df <- data.frame(id = 1:6, value = c(3, 8, 999, 2, 5, 7))
reduce_rows(df, function(d) any(d$value > 100))
```

The result is usually small enough to capture with `dput()`.

## The algorithm

Both `minex()` and `reduce_rows()` are front ends to `ddmin()`. The default
strategy is Counter-Based Delta Debugging (CDD; Zhang et al., 2025), which
skips deletion attempts that its counters identify as redundant. The classic
block-halving algorithm of Zeller and Hildebrandt (2002) remains available with
`algorithm = "ddmin"`. `ddmin()` works on any collection plus a predicate, so it
is reusable on its own.

```{r}
ddmin(strsplit("the quick brown fox", " ")[[1]], function(s) "fox" %in% s)
```

After its main reduction phase, `ddmin()` repeatedly sweeps the remaining
elements and removes any single element whose deletion preserves the behavior.
This fixed-point verification step (Vince and Kiss, 2024) makes the output
*one-minimal*: removing any single element makes the behavior disappear.

At the default granularity, `minex()` first works over top-level statements. If
that pass removes nothing, it automatically retries within the surviving
statements using HDD* (Hierarchical Delta Debugging; Misherghi and Su, 2006).
The expression pass can remove pipeline stages (`|>` and `%>%`) and positional
call arguments. You can request it directly with
`granularity = "expression"`, or suppress the automatic retry by explicitly
passing `granularity = "statement"`.

## Limitations

`minex()` targets R-level conditions (errors). Failures that crash the R process
or hang are not captured as conditions; with the `callr` backend they are simply
treated as not reproducing the target. Expression-level reduction is currently
limited to pipeline stages and positional call arguments; it does not rewrite
arbitrary R syntax.

## References

- Misherghi, G. and Su, Z. (2006). "HDD: Hierarchical Delta Debugging."
  *ICSE 2006*, 142-151. <https://doi.org/10.1145/1134285.1134307>
- Vince, D. and Kiss, A. (2024). "Evaluation of the Fixed-Point Iteration of
  Minimizing Delta Debugging." *Journal of Software: Evolution and Process*,
  36(10), e2702. <https://doi.org/10.1002/smr.2702>
- Zeller, A. and Hildebrandt, R. (2002). "Simplifying and Isolating
  Failure-Inducing Input." *IEEE Transactions on Software Engineering*, 28(2),
  183-200. <https://doi.org/10.1109/32.988498>
- Zhang, M., Xu, Z., Tian, Y., Cheng, X. and Sun, C. (2025). "Toward a Better
  Understanding of Probabilistic Delta Debugging." *ICSE 2025*.
  <https://doi.org/10.1109/ICSE55347.2025.00117>
